Safety Cases for Software Product Lines

### 2017-10-04 doc/SafetyCases/SafetyCasesPL.tex
Definitions with FLDPL examples

%\subsection{Running Example}
%
%Throughout this document we use a Fuel Level Display Product Line (FLDPL) example from \cite{Gallucci_2013}. This is a simply product line safety case that extends a single product ISO26262 safety case \cite{Dardar_2014}. The diagrams in this document also come from \cite{Gallucci_2013} unless otherwise specified.
%
%\section{Existing Techniques}
%
\section{Software Product Lines (SPLs)}

\subsection{Basic Concepts}

@@

@@

For example, let us consider a product line of simple object stores. An object store can support the storage of a single object or multiple objects. Orthogonally, the store can provide some access control mechanism allowing only certain users to store objects. Store capacity (SingleObject vs MultiObject) and access control are just examples of possible variability points in an SPL.

@@

In our object store SPL, SingleObject, MultiObject and access control are the 3 features within this product line.

An individual product can be generated from an SPL by selecting a set of features for that product. Not all product combinations are valid though. For example, it does not make sense for an object store to be SingleObject and MultiObject at the same time. Those two features are supposed to be mutually exclusive in our product line. This kind of constraints are captured by a \emph{Feature Model}:

@@

For example, a boolean formula expressing our mutual exclusion condition for SingleObject and MultiObject stores, together with an optional AccessControl feature, would be:

$(SingleObject \vee MultiObject) \wedge (\neg SingleObject \vee \neg MultiObject) $

@@

@@

For example, a product where we have the MultiObject and AccessControl features can be represented by the following PC: $ MultiObject \wedge AccessControl $

\subsection{SPL Engineering Approaches}

@@

\subsubsection{Generative}

@@

\subsubsection{Annotative}

@@

\subsubsection{Delta-Oriented}

>>>>

\section{Running Example}

Throughout this document we use a Fuel Level Display Product Line (FLDPL) example from \cite{Gallucci_2013}. This is a simply product line safety case that extends a single product ISO26262 safety case \cite{Dardar_2014}. The diagrams in this document also come from \cite{Gallucci_2013} unless otherwise specified.

The Fuel Level Display (FLD) system is an automotive component implementing two functionalities:

\begin{enumerate}
\item Displaying the fuel level of a fuel tank on the driver's dashboard
\item Triggering a warning when the fuel level goes below a certain threshold (optional feature)
\end{enumerate}

The Fuel Level Display Product Line (FLDPL) captures four variants of FLDs based on the vehicle type and fuel type:

\begin{enumerate}
\item Truck with liquid fuel
\item Bus with liquid fuel
\item Truck with gas
\item Bus with gas
\end{enumerate}

section{Software Product Lines (SPLs)}

@@

\subsection{Software Product Line}

@@

For our running example, we have four variants of Fuel Level Display systems. In addition, each of those variants can come with or without the warning feature. This family of product variants has a lot in common, so it is more cost effective to build them as an SPL instead of building each product in isolation.

\subsection{Features}

@@

At a high level, our FLDPL has two features: fuel level display and low fuel level warning.

Orthogonally, variation with respect to vehicle types and fuel types are also considered feature variants.

So this is the FLDPL list of features:

\begin{enumerate}
\item FuelLevel
\item FuelWarning (not present in trucks with gas)
\item Truck
\item Bus
\item LiquidFuel
\item Gas
\end{enumerate}

\subsection{Feature Model}

@@

For example, a boolean formula expressing the mutual exclusion condition for the Truck and Bus features:

$(Truck \vee Bus) \wedge (\neg Truck \vee \neg Bus) $

\subsection{Feature Diagram}

@@

\subsection{Presence Condition}

@@

For example, an FLD product for buses, with liquid fuel, with fuel level display and low fuel warning can be represented by the following PC:

Bus $\wedge$ Liquid $\wedge$ Estimation $\wedge$ Warning

\section{SPL Engineering Approaches}

@@

\subsection{Generative}

@@

\subsection{Annotative}

@@

\subsection{Delta-Oriented}
